A geometric approach to linear and quadratic Chabauty

### 2020-06-11 MS.tex
Fixed most of the last comments by Bas

%-----------------------------------------------------------------------------------------------------------------------
\todo{Remove overfull h-box warnings}\todo{Referentie naar een artikel van Balakrishnan en Dogra, en Besser.}
\section{Introduction}
\label{section:intro}
In mathematics a fundamental problem is solving polynomial equations over the rationals, dating back to Diophantus. An important special case in algebraic geometry is that of curves. The behavior of the rational points of curves depends enormously on the \textit{genus} $g$ of the curve $C$, a numerical invariant of a curve. For $g = 0$, there are either no or infinitely many solutions, and their behavior is well understood. For $g = 1$, we get either no point or an elliptic curve, whose set of rational points forms an often infinite group.

@@

Before it was proven, one of the major partial results was a theorem from Chabauty in 1941, phrased in terms of the rank $r$ of the group of rational points on the Jacobian $J$ of the curve, also called the Mordell-Weil rank, which is finite by the Mordell-Weil theorem. Chabauty proved, using $p$-adic methods, that if $r$ is strictly smaller than the genus $g$, then $C$ has only finitely many rational points. Loosely said, this proof and all Chabauty related theorems, rely on choosing a prime $p$ for which $C$ has good reduction, and intersecting $J(\Z)$ with $C(\Z_p)$ inside the bigger Lie group $J(\Z_p)$; by properties of $\Z_p$, the subgroup $J(\Z)$, generated by $r$ elements, lies within a Lie subgroup of dimension at most $r$, and $C(\Z_p)$ is a $1$-dimensional $p$-adic manifold, and hence their intersection can be proven to be discrete and -- as $J(\Z_p)$ is compact -- finite.

@@

This Coleman-Chabauty method has been greatly generalised by Kim to so called non-abelian Chabauty. He interprets working in the Jacobian as dealing with the abelianised fundamental group of the curve $C$, and works with larger, non-abelian quotients of the fundamental group. Using quadratic Chabauty, the simplest case of non-abelian Chabauty, Balakrishnan, Dogra, M\"uller, Tuitman and Vonk in 2017 famously calculated  all rational points of the ``cursed curve'', the modular curve $X_s^+(13)$ \cite{cursedcurve}. They use an endomorphism of the Jacobian, and do $p$-adic analysis on $p$-adic local heights to make quadratic Chabauty explicit.

Very recently, Edixhoven and Lido made available a preprint of their article ``Geometric quadratic Chabauty'' \cite{edixhoven20}. Their goal in this article is to make a more geometric approach to effective quadratic Chabauty, by working in a pullback $T$ of the Poincar\'e torsor of the Jacobian. They also abandon the Coleman method for Chabauty; instead they parametrise the map $T(\Z) \to T(\Z_p)$ with power series, and pull back equations for the curve inside $T(\Z_p)$ along this map.

The purpose of this thesis is to make the work by Edixhoven and Lido more accessible by applying geometric Chabauty to the linear case, i.e. when working with $J(\Z)$ and $J(\Z_p)$. In this thesis, we go into more detail of what is happening throughout the process, including both theoretical and practical components. This new method of applying Chabauty leads to, when working modulo $p$, a conditional upper bound $|C(\Q)| \leq |C(\F_p)|$, as obtained in Proposition~\ref{prop:linchabauty}. This upper bound compares favorably with classical Coleman-Chabauty, but unlike classical Coleman-Chabauty, we cannot say in general when the conditions of these propositions hold. We develop heuristics that suggest that generally trying multiple primes $p$ of good reduction will yield good results; these heuristics say that the conditional upper bound will hold for a set of primes of density $1$, and if $r < g-1$ can even often (in a set of primes of density $e^{-1}$) be improved to prove there are at most $|C(\Q)|$ rational points\todo{Bas, dit is denk ik de beste samenvatting}. The methods used to obtain Proposition~\ref{prop:linchabauty} are also amenable to higher precision calculations. We also give a hyperelliptic example with $r = 1, g = 2$, where the bound from Proposition~\ref{prop:linchabauty} indeed holds.

\subsection{Overview}
We first present the context in which we will perform Chabauty. Let $p>2$ be a prime, and $C$ be a scheme over $\Z$, proper, flat, regular with $C$ smooth over $\Z_{(p)}$ and $C_\Q$ of dimension $1$ and geometrically connected, of genus $g \geq 1$ and with the Jacobian $J$ of $C$ having Mordell-Weil rank $r < g$. Assume we have a $\Q$-point $b$ in $C$, or equivalently a $\Z$-point. We view $C$ as a subscheme of $J$, using the map $Q \mapsto Q - b$ on points. Let $P \in C(\F_p)$ be a point such that $t := P - b \in J(\F_p)$ lies in the image of $J(\Z)$.

@@

Geometric chabauty works by finding an upper bound for the cardinality of $C(\Z)_P$. For this, we use the following diagram.
\[
\begin{tikzcd}
C(\Z)_P \arrow[d] \arrow[r, "a"] & J(\Z)_t \arrow[d] \\
C(\Z_p)_P \arrow[r]              & J(\Z_p)_t        
\end{tikzcd}
\]
where the two vertical maps are inclusions, and the two horizontal maps are subtraction of $b$. We will in fact compute upper bounds for the larger set $C(\Z_p)_P \cap \overline{J(\Z)_t}$, where $\overline{J(\Z)_t}$ is the closure of $J(\Z)_t$ in $J(\Z_p)_t$.

@@

\subsection{Preliminaries}\todo{Bas, dit en het volgende stukje zijn nieuw}
We expect the readers of this thesis to be familiar with algebraic geometry, e.g. master students with a specialisation in algebraic geometry. The introduction to geometric quadratic Chabauty in Section~\ref{section:quadratic} contains some terms typically not treated in a master; while knowing them will help put the material in context, this section is meant to be understandible even if that is not the case. \todo{Bas, nog een specifieke referentie voor voorkennis die ik op kan nemen, of niet per se?}

@@

\todo{Volgend stukje is nieuw, en het stukje erna is iets herschreven}\begin{rem}
\label{rem:torsortangentspace}
In actual calculations, we will be focusing on $X(\Z/p^2\Z)_x$. This set is a natural torsor of the tangent space $T_x X_{\F_p}$ of $X_{\F_p}$ at $x$, which we will describe here. We write $R$ for the local ring $O_{X_{\Z/p^2\Z},x}$ and $\m$ for its maximal ideal, and $\overline{R}$ and $\overline{m}$ for their reductions modulo $p$. Then an element in $X(\Z/p^2\Z)_x$ is a local morphism $R \to \Z/p^2\Z$. As $R/\m = \F_p$, giving such a local morphism is equivalent to giving a map $\m \to p\Z/p^2\Z$ respecting multiplication and sending $p$ to $p$. Such a map factors uniquely through $\m/\m^2 \to p\Z/p^2\Z$. We see the set $X(\Z/p^2\Z)_x$ is canonically in bijection with $\F_p$-linear maps $\m/\m^2 \to p\Z/p^2\Z$ sending $p$ to $p$. Note that $\m/\m^2$ is a $d+1$-dimensional $\F_p$-vector space, and $\overline{\m}/\overline{\m}^2$ is a $d$-dimensional vector space; the map $\m/\m^2 \to \overline{\m}/\overline{\m}^2$ is dividing out by $p$. Denoting ${}^\wedge$ to be $\Hom(\cdot,\F_p)$, the exact sequence
\[
0 \to p\Z/p^2\Z \to \m/\m^2 \to \overline{\m}/\overline{\m}^2 \to 0
\]
can be dualised to
\[
0 \to (\overline{\m}/\overline{\m}^2)^\wedge \to (\m/\m^2)^\wedge \to (p\Z/p^2\Z)^{\wedge} \to 0.
\]
As $X(\Z/p^2\Z)_x$ is exactly the subset of $(\m/\m^2)^\wedge$ that gets mapped to $p \mapsto 1 \in (p\Z/p^2\Z)^{\wedge}$, we see this is naturally a torsor under $(\overline{\m}/\overline{\m}^2)^\wedge = T_{x} X_{\F_p}$.

Note that this set does not have any more canonical structure; for example, with $X = \A^1_{\Z_p}$ and $x$ is the $\F_p$-point $1$, the set $X(\Z/p^2\Z)_x$ is $\{1 + pi | i \in \F_p\}$, and we see we cannot upgrade the torsor structure. This is due to the fact that even after choosing a parameter at $x$ as a point of $X_{\F_p}$, a lift of such a parameter to $X_{\Z/p^2\Z}$ is not unique, and can in fact differ up to translation. This is in contrast to the situation over $\F_p[\varepsilon]/(\varepsilon^2)$, where the $\F_p$-algebra structure gives rise to an isomorphism of $\F_p$-vector spaces between $X(\F_p[\varepsilon]/(\varepsilon^2))$ and $T_x X_{\F_p}$; indeed, this is an alternate definition of the tangent space.
\end{rem}

@@

\subsection{From group schemes to formal groups}
We first recall the definition of a formal group. For this entire section, $R$ is a ring, and $x,y,z$ are vectors of variables.
\begin{defn}
\label{defn:formalgroup}
Let $n$ be a non-negative integer. Let $x,y,z$ be vectors of $n$ variables. An $n$-dimensional formal group is an element $F = (F_1,\dots,F_n)$ of $R[[x,y]]_0^n$ satisfying $F \equiv x + y \bmod (x,y)^2$ and $F(F(x,y),z) = F(x,F(y,z))$. If furthermore $F(x,y) = F(y,x)$, this formal group is said to be commutative.
\end{defn}
\begin{exmp}
\label{exmp:ganformalgroup}
For any $n$, we can take $F(x,y) = x+ y$, also known as the $n$-dimensional additive formal group.
\end{exmp}
\begin{exmp}
\label{exmp:gmformalgroup}
Take $n=1$ and $F(x,y) = x + y + xy = (1+x)(1+y)-1$, also known as the multiplicative formal group, as it is a translation of the natural multiplication on $1 + xR[x]$. This satisfies associativity by the formula \[F(F(x,y),z) = (1+x)(1+y)(1+z)-1 = F(x,F(y,z)).\]
\end{exmp}
Note there is no mention of an inverse; the following lemma, a formal version of the implicit function theorem, tells us that the inverse follows automatically from the definitions.
\begin{lem}
\label{lem:implicitfunction}
Let $x,y$ be vectors of variables of length $n$. Let $F \in R[[x,y]]_0^n$ such that $F \equiv Ax + By \bmod (x,y)^2$ with $A \in \Mat_n(R)$ and $B \in \GL_n(R)$. Then there is a unique $\iota \in R[[x]]_0^n$ such that $F(x,\iota(x)) = 0$. 
\end{lem}
\begin{proof}
If $F$ is a polynomial, this is a special case of a multivariate version of Hensel's lemma, as in Corollaire~2 of \cite[III,4.5]{BourbakiCA}: we work over $R[[x]]$ which is complete with respect to the ideal $\m = (x)$, and the derivative matrix of $F(x,\iota)$ with respect to $\iota$ is $B$, which is invertible, and $\iota = 0$ gives a solution modulo $\m$. Now for $F$ any power series, let $F_j$ be the polynomial consisting of all terms in $F$ of degree at most $j$, and let $\iota_j$ be the unique power series in $R[[x]]_0^n$ such that $F_j(x,\iota_j) = 0$. Note that both $\iota_j$ and $\iota_{j+1}$ are solutions to $F_j(x,\iota) \equiv 0 \bmod \m^j$, so by the uniquess guaranteed by Hensel's lemma used over $R[[x]]/\m^j$, these must be equal. Hence they converge to $\iota \in R[[x]]_0^n$, which is the unique solution of $F(x,\iota(x)) = 0$.
\end{proof}
This has the following corollary about the inverse of a power series.
\begin{cor}
\label{lem:formalinverse}
Let $x$ have length $n$. Let $a \in R[[x]]_0^n$ with $a \equiv Px \bmod (x)^2$ for some matrix $P\in \GL_n(R)$. Then there is a unique $b \in R[[x]]_0^n$ such that $a \circ b = b \circ a = x$.
\end{cor}
\begin{proof}
Let $F(x,y) = x - a(y)$. This satisfies the constraints of Lemma~\ref{lem:implicitfunction}, so we find a unique $b$ such that $a \circ b = x$. Applying Lemma~\ref{lem:implicitfunction} again gives a unique $c$ such that $b \circ c = x$. But then $a = a \circ (b \circ c) = (a \circ b) \circ c = c$ shows $a = c$ and hence we are done.
\end{proof}
So a formal group $F$ does by Lemma~\ref{lem:implicitfunction} indeed have a right inverse $\iota_F$. Also, by $F(0,0) = 0$ we have that $F(x,F(\iota_f,0)) = 0$ so $F(\iota_f,0)$ is in fact equal to $\iota_F$; as $\iota_F \equiv -x \bmod (x)^2$, it has a formal inverse and hence $F(x,0) = x$, so $0$ is indeed a right unit. A standard argument now shows that $\iota_F$ is also a left inverse and $0$ is also a right unit, so $F$ gives $R[[x]]$, together with the $(x)$-adic topology, the structure of a group object in the category of topological $R$-algebras\todo{Nu zou het goed moeten zijn, toch? Als nog steeds niet: werk in de categorie $(S,I)$ waar $S$ compleet is m.r.t. $I$}; for two continuous morphisms $f_1,f_2$ from $R[[x]]$ to some ring $A$, given by sending $x$ to $a_1,a_2$ respectively, we define $(f_1 \oplus f_2)(x) = F(a_1,a_2)$, and by our considerations this makes $\Hom_{\text{cont}}(R[[x]],A)$ into a group, functorially in $A$.

Given a smooth scheme $G$ over $\Z_p$ of relative dimension $d$, together with a $\Z_p$-point $e$, we can look at the completion $\widehat{O}_{G,e}$ along the section $e$. By smoothness, after picking parameters, this is isomorphic as topological $\Z_p$-algebras to the ring of power series $\Z_p[[x_1,\dots,x_d]]$. This completion naturally gives rise to a formal scheme denoted $\Spf \widehat{O}_{G,e}$; in a categorical notion, this is a functor on finite length $\Z_p$-algebras sending $A$ to $\Hom_{\text{cont}}(\widehat{O}_{G,e},A)$, but we can also think of it as a locally ringed space with $\Spec \Z_p$ as topological space, and sheaf of rings $\lim_n \widetilde{O_{G,e}/m^n}$ where $m$ is the ideal of $O_{G,e}$ defining $e$\todo{Is dit de juiste formulering?}. The global sections of this sheaf are simply $\widehat{O}_{G,e}$.

@@

This leads to a now trivial proposition.\todo{Bas, proposition of theorem?}
\begin{prop}
\label{prop:linchabauty}
If for all points $P \in C(\F_p)$, the linear system of $g-1$ equations modulo $p$ defining $C(\Z_p)_P \subset J(\Z_p)_t$ pulled back to $\overline{J(\Z)_t}$ has at most one solution, we get the inequality
\[
|C(\Z)| \leq |C(\F_p)|.
\]
\end{prop}

@@

\section{Complications and improvements}
\label{section:remarks} \todo{Bas, het is niet alleen werk van anderen, hopelijk is het nu duidelijker}
In this section, we quickly make some further remarks on how or when to apply the theory we have built up in previous section to real cases.
\begin{itemize} 
\item Even if $r \geq g$, it might still be possible that $\overline{J(\Z)}$ coincidentally has dimension $r'$ smaller than $r$ (for example, if $r > g$); then one can find a rank $r'$ subgroup of $J(\Z)$ with the same closure inside $J(\Z_p)$, and do computations in this subgroup; we can expect this to work if $r' < g$. We can even use this to improve our calculations if we already have $r < g$, but $r'$ is even smaller. On the other hand, if $J(\Z)$ is dense in $J(\Z_p)$, there is no way to apply Chabauty's method to the curve.
\item Note that in all of the discussion of the previous section, we are calculating upper bounds for the cardinality of $C(\Z_p)_P \cap \overline{J(\Z)_t}$. It can and does however happen that this set is strictly bigger than $C(\Z)_P$, as shown in \cite{Balakrishnan19}, where Coleman-Chabauty is used on a whole database of curves. They treat several possible cases and examples where the set is bigger. For example, there might be a $K$-point $Q$ where $K$ is a quadratic field where $p$ splits, such that $Q-b$ lies in $J(\Z_p)_0$ and is non-$p$-torsion in $J(\Z_p)_0/J(\Z)_0$; then, as $J(\Z_p)$ is isomorphic as continous group to $\Z_p^g$, the point $Q-b$ lies in $\overline{J(\Z)_0}$, and clearly also in $C(\Z_p)$.\todo{Bas, explanation added}
\item An important part of doing these calculations, is working with the Mordell-Weil group $J(\Z)$. In general, finding generators of $J(\Z)$ is computationally out of reach, and one may need to assume the BSD conjecture to even find the rank $r$. However, note that one does not need to find all of $J(\Z)$; if we can generate a subgroup of $J(\Z)$ of index finite and coprime to $p|J(\F_p)|$, it will have the same closure in $J(\Z_p)$. For this, we only need to generate sufficiently many independent points of the Jacobian, and saturate subgroups with respect to some primes. Both of these feats are easier than finding all of $J(\Z)$.
\item Even this method does not work for a certain prime $p$, and neither for some other primes we tried, we can still use parts of the information we have learned. For example, maybe the method tells us that we have found all points of $J(\Z)$ stemming from $C(\Z)$ except those in a certain fibre of the map $J(\Z) \to J(\F_p)$. We can then aggregate this information for different primes, by looking at the maps $J(\Z) \to \prod_{p \in S} J(\F_p)$ for some set $S$ of primes of good reduction. See for example \cite{stoll10} for an introduction to the Mordell-Weil sieve.
\end{itemize}

@@

Code for the calculations in this section can be found in \cite{spelier2020}.
\todo{Bas, hiervoor ben ik nu gegaan}

>>>>

%-----------------------------------------------------------------------------------------------------------------------
\todo{Remove overfull h-box warnings}
\section{Introduction}
\label{section:intro}
In mathematics a fundamental problem is solving polynomial equations over the rationals, dating back to Diophantus. An important special case in algebraic geometry is that of curves. The behavior of the rational points of curves depends enormously on the \textit{genus} $g$ of the curve $C$, a numerical invariant of a curve. For $g = 0$, there are either no or infinitely many solutions, and their behavior is well understood. For $g = 1$, we get either no point or an elliptic curve, whose set of rational points forms an often infinite group.

@@

Before it was proven, one of the major partial results was a theorem from Chabauty in 1941, phrased in terms of the rank $r$ of the group of rational points on the Jacobian $J$ of the curve, also called the Mordell-Weil rank, which is finite by the Mordell-Weil theorem. Chabauty proved, using $p$-adic methods, that if $r$ is strictly smaller than the genus $g$, then $C$ has only finitely many rational points. Loosely said, this proof and all Chabauty related theorems, rely on choosing a prime $p$ for which $C$ has good reduction, and intersecting $J(\Z)$ with $C(\Z_p)$ inside the bigger Lie group $J(\Z_p)$; by properties of $\Z_p$, the subgroup $J(\Z)$, up to torsion generated by $r$ elements, lies within a Lie subgroup of dimension at most $r$, and $C(\Z_p)$ is a $1$-dimensional $p$-adic manifold, and hence their intersection can be proven to be discrete and -- as $J(\Z_p)$ is compact -- finite.\todo{Ik moet volgens mij uberhaupt nadenken wat over $\Q,\Q_p,\Z,\Z_p$ moet}

@@

This Coleman-Chabauty method has been greatly generalised by Kim to so called non-abelian Chabauty in \todo{Ref} and \todo{Ref}. He interprets working in the Jacobian as dealing with the abelianised fundamental group of the curve $C$, and works with larger, non-abelian quotients of the fundamental group. Then the simplest case of non-abelian Chabauty, quadratic Chabauty was developed by Kim, Balakrishnan, Besser Dogran and M\"uller in the series of articles \todo{Ref}, and further extended by Balakrishnan, Dogra, M\"uller, Tuitman and Vonk; in 2017 they famously calculated  all rational points of the ``cursed curve'', the modular curve $X_s^+(13)$ \cite{cursedcurve}. They use an endomorphism of the Jacobian, and do $p$-adic analysis on $p$-adic local heights to make quadratic Chabauty explicit. Their methods are strong enough to prove finiteness of $C(\Z)$ under the condition of $r < g + \rho - 1$, where $\rho$ is the N\'eron-Severi rank of the Jacobian.

Very recently, Edixhoven and Lido made available a preprint of their article ``Geometric quadratic Chabauty'' \cite{edixhoven20}. Their goal in this article is to make a more geometric approach to effective quadratic Chabauty, by working in a pullback $T$ of the Poincar\'e torsor of the Jacobian. They also abandon the Coleman method for Chabauty; instead they parametrise the map $T(\Z) \to T(\Z_p)$ with power series, and pull back equations for the curve inside $T(\Z_p)$ along this map. Again, a necessary condition is $r < g + \rho - 1$.

The purpose of this thesis is to make the work by Edixhoven and Lido more accessible by applying geometric Chabauty to the linear case, i.e. when working with $J(\Z)$ and $J(\Z_p)$ and assuming $r < g$. In this thesis, we go into more detail of what is happening throughout the process, including both theoretical and practical components. This new method of applying Chabauty leads to, when working with single digit $p$-adic precision, a conditional upper bound $|C(\Q)| \leq |C(\F_p)|$, as obtained in Proposition~\ref{prop:linchabauty}. This upper bound compares favorably with classical Coleman-Chabauty, but unlike classical Coleman-Chabauty, we cannot say in general when the conditions of these propositions hold. We develop heuristics that suggest that generally trying multiple primes $p$ of good reduction will yield good results; these heuristics say that the conditional upper bound will hold for a set of primes of density $1$, and if $r < g-1$ can even often (in a set of primes of density at least $e^{-1}$) be improved to prove there are at most $|C(\Q)|$ rational points. The methods used to obtain Proposition~\ref{prop:linchabauty} are also amenable to higher precision calculations; indeed, in the case $r = g-1$ this may be necessary to calculate $|C(\Z)|$ exactly.\todo{Over hebben met Bas}

We also give a hyperelliptic example with $r = 1, g = 2$, where the bound from Proposition~\ref{prop:linchabauty} indeed holds, and we are able to compute $C(\Z)$.

\subsection{Overview}
We first present the context in which we will perform Chabauty. Let $p>2$ be a prime, and $C$ be a scheme over $\Z$, proper, flat, regular with $C$ smooth over $\Z_{(p)}$ and $C_\Q$ of dimension $1$ and geometrically connected, of genus $g \geq 1$ and with the N\'eron model $J$ of the Jacobiano of $C_\Q$ having Mordell-Weil rank $r < g$. Assume we have a $\Q$-point $b$ in $C$, or equivalently a $\Z$-point. We view $C$ as a subscheme of $J$, using the map $Q \mapsto Q - b$ on points. Let $P \in C(\F_p)$ be a point such that $t := P - b \in J(\F_p)$ lies in the image of $J(\Z)$.

@@

Geometric Chabauty works by finding an upper bound for the cardinality of $C(\Z)_P$. For this, we use the following diagram.
\[
\begin{tikzcd}
C(\Z)_P \arrow[d] \arrow[r, "a"] & J(\Z)_t \arrow[d] \\
C(\Z_p)_P \arrow[r]              & J(\Z_p)_t        
\end{tikzcd}
\]
where the two vertical maps are inclusions, and the two horizontal maps are subtraction of $b$. We will in fact compute upper bounds for the larger set $C(\Z_p)_P \cap \overline{J(\Z)_t}$, where $\overline{J(\Z)_t}$ is the closure of $J(\Z)_t$ in $J(\Z_p)_t$.

@@

\subsection{Preliminaries}
We expect the readers of this thesis to be familiar with algebraic geometry, e.g. master students with a specialisation in algebraic geometry. The introduction to geometric quadratic Chabauty in Section~\ref{section:quadratic} contains some terms typically not treated in a master; while knowing them will help put the material in context, this section is meant to be understandible even if that is not the case.

@@

\begin{rem}
\label{rem:torsortangentspace}
In actual calculations, we will be focusing on $X(\Z/p^2\Z)_x$. This set is a natural torsor of the tangent space $T_x X_{\F_p}$ of $X_{\F_p}$ at $x$, which we will describe here. We write $R$ for the local ring $O_{X_{\Z/p^2\Z},x}$ and $\m$ for its maximal ideal, and $\overline{R}$ and $\overline{m}$ for their reductions modulo $p$. Then an element in $X(\Z/p^2\Z)_x$ is a local morphism $R \to \Z/p^2\Z$. As $R/\m = \F_p$, giving such a local morphism is equivalent to giving a map $\m \to p\Z/p^2\Z$ respecting multiplication and sending $p$ to $p$. Such a map factors uniquely through $\m/\m^2 \to p\Z/p^2\Z$. We see the set $X(\Z/p^2\Z)_x$ is canonically in bijection with $\F_p$-linear maps $\m/\m^2 \to p\Z/p^2\Z$ sending $p$ to $p$. Note that $\m/\m^2$ is a $d+1$-dimensional $\F_p$-vector space, and $\overline{\m}/\overline{\m}^2$ is a $d$-dimensional vector space; the map $\m/\m^2 \to \overline{\m}/\overline{\m}^2$ is dividing out by $p$. Denoting ${}^\wedge$ to be $\Hom(\cdot,\F_p)$, the exact sequence
\[
0 \to p\Z/p^2\Z \to \m/\m^2 \to \overline{\m}/\overline{\m}^2 \to 0
\]
can be dualised to
\[
0 \to (\overline{\m}/\overline{\m}^2)^\wedge \to (\m/\m^2)^\wedge \to (p\Z/p^2\Z)^{\wedge} \to 0.
\]
As $X(\Z/p^2\Z)_x$ is exactly the subset of $(\m/\m^2)^\wedge$ that gets mapped to $p \mapsto 1 \in (p\Z/p^2\Z)^{\wedge}$, we see this is naturally a torsor under $(\overline{\m}/\overline{\m}^2)^\wedge = T_{x} X_{\F_p}$.

Note that this set does not have any more canonical structure; for example, with $X = \A^1_{\Z_p}$ and $x$ is the $\F_p$-point $1$, the set $X(\Z/p^2\Z)_x$ is $\{1 + pi | i \in \F_p\}$, and we see we cannot upgrade the torsor structure. This is due to the fact that even after choosing a parameter at $x$ as a point of $X_{\F_p}$, a lift of such a parameter to $X_{\Z/p^2\Z}$ is not canonical, and can in fact differ up to translation. This is in contrast to the situation over $\F_p[\varepsilon]/(\varepsilon^2)$, where the $\F_p$-algebra structure gives rise to an isomorphism of $\F_p$-vector spaces between $X(\F_p[\varepsilon]/(\varepsilon^2))$ and $T_x X_{\F_p}$; indeed, this is an alternate definition of the tangent space.
\end{rem}

@@

\subsection{From group schemes to formal groups}
We first recall the definition of a formal group. For this entire section, $R$ is a ring, and $x,y,z$ are vectors of variables.
\begin{defn}
\label{defn:formalgroup}
Let $n$ be a non-negative integer. Let $x,y,z$ be vectors of $n$ variables. An $n$-dimensional formal group is an element $F = (F_1,\dots,F_n)$ of $R[[x,y]]_0^n$ satisfying $F \equiv x + y \bmod (x,y)^2$ and $F(F(x,y),z) = F(x,F(y,z))$. If furthermore $F(x,y) = F(y,x)$, this formal group is said to be commutative.
\end{defn}
\begin{exmp}
\label{exmp:ganformalgroup}
For any $n$, we can take $F(x,y) = x+ y$, also known as the $n$-dimensional additive formal group.
\end{exmp}
\begin{exmp}
\label{exmp:gmformalgroup}
Take $n=1$ and $F(x,y) = x + y + xy = (1+x)(1+y)-1$, also known as the multiplicative formal group, as it is a translation of the natural multiplication on $1 + xR[x]$. This satisfies associativity by the formula \[F(F(x,y),z) = (1+x)(1+y)(1+z)-1 = F(x,F(y,z)).\]
\end{exmp}
Note there is no mention of an inverse; the following lemma, a formal version of the implicit function theorem, tells us that the inverse follows automatically from the definitions.
\begin{lem}
\label{lem:implicitfunction}
Let $x,y$ be vectors of variables of length $n$. Let $F \in R[[x,y]]_0^n$ such that $F \equiv Ax + By \bmod (x,y)^2$ with $A \in \Mat_n(R)$ and $B \in \GL_n(R)$. Then there is a unique $\iota \in R[[x]]_0^n$ such that $F(x,\iota(x)) = 0$. 
\end{lem}
\begin{proof}
If $F$ is a polynomial, this is a special case of a multivariate version of Hensel's lemma, as in Corollaire~2 of \cite[III,4.5]{BourbakiCA}: we work over $R[[x]]$ which is complete with respect to the ideal $\m = (x)$, and the derivative matrix of $F(x,\iota)$ with respect to $\iota$ is $B$, which is invertible, and $\iota = 0$ gives a solution modulo $\m$. Now for $F$ any power series, let $F_j$ be the polynomial consisting of all terms in $F$ of degree at most $j$, and let $\iota_j$ be the unique power series in $R[[x]]_0^n$ such that $F_j(x,\iota_j) = 0$. Note that both $\iota_j$ and $\iota_{j+1}$ are solutions to $F_j(x,\iota) \equiv 0 \bmod \m^j$, so by the uniquess guaranteed by Hensel's lemma used over $R[[x]]/\m^j$, these must be equal. Hence they converge to $\iota \in R[[x]]_0^n$, which is the unique solution of $F(x,\iota(x)) = 0$.
\end{proof}
This has the following corollary about the inverse of a power series.
\begin{cor}
\label{lem:formalinverse}
Let $x$ have length $n$. Let $a \in R[[x]]_0^n$ with $a \equiv Px \bmod (x)^2$ for some matrix $P\in \GL_n(R)$. Then there is a unique $b \in R[[x]]_0^n$ such that $a \circ b = b \circ a = x$.
\end{cor}
\begin{proof}
Let $F(x,y) = x - a(y)$. This satisfies the constraints of Lemma~\ref{lem:implicitfunction}, so we find a unique $b$ such that $a \circ b = x$. Applying Lemma~\ref{lem:implicitfunction} again gives a unique $c$ such that $b \circ c = x$. But then $a = a \circ (b \circ c) = (a \circ b) \circ c = c$ shows $a = c$ and hence we are done.
\end{proof}
So a formal group $F$ does by Lemma~\ref{lem:implicitfunction} indeed have a right inverse $\iota_F$. Also, by $F(0,0) = 0$ we have that $F(x,F(\iota_f,0)) = 0$ so $F(\iota_f,0)$ is in fact equal to $\iota_F$; as $\iota_F \equiv -x \bmod (x)^2$, it has a formal inverse and hence $F(x,0) = x$, so $0$ is indeed a right unit. A standard argument now shows that $\iota_F$ is also a left inverse and $0$ is also a right unit, so $F$ gives $(R[[x]],(x))$, together with the $(x)$-adic topology, the structure of a group object in the category of pairs of objects $(A,I)$ where $A$ is a topological $R$-algebra, complete with respect to $I$ with the $I$-adic topology; for two continuous morphisms $f_1,f_2$ from $(R[[x]],(x))$ to $(A,I)$, given by sending $x$ to $a_1,a_2 \in I^n$ respectively, we define $(f_1 \oplus f_2)(x) = F(a_1,a_2)$, and by our considerations this makes $\Hom_((R[[x]],x),(A,I))$ into a group, functorially in $A$.\todo{Nu zou het goed moeten zijn, toch?}

Given a smooth scheme $G$ over $\Z_p$ of relative dimension $d$, together with a $\Z_p$-point $e$, we can look at the completion $\widehat{O}_{G,e}$ along the section $e$. By smoothness, after picking parameters, this is isomorphic as topological $\Z_p$-algebras to the ring of power series $\Z_p[[x_1,\dots,x_d]]$. This completion naturally gives rise to a formal scheme denoted $\Spf \widehat{O}_{G,e}$; in a categorical notion, this is a functor on finite length $\Z_p$-algebras sending $A$ to $\Hom_{\text{cont}}(\widehat{O}_{G,e},A)$, but we can also think of it as a locally ringed space with $\Spec \Z_p$ as topological space, and sheaf of rings $\lim_n \widetilde{O_{G,e}/m^n}$ where $m$ is the ideal of $O_{G,e}$ defining $e$\todo{Over hebben met Bas}. The global sections of this sheaf are simply $\widehat{O}_{G,e}$.

@@

This leads to a now trivial proposition.
\begin{prop}
\label{prop:linchabauty}
If for all points $P \in C(\F_p)$, the linear system of $g-1$ equations modulo $p$ defining $C(\Z_p)_P \subset J(\Z_p)_t$ pulled back to $\overline{J(\Z)_t}$ has $n_P \leq 1$ solutions, we get the inequality
\[
|C(\Z)| \leq \sum_{P \in C(\F_p)} n_P \leq |C(\F_p)|. 
\]\todo{Over hebben met Bas}
\end{prop}

@@

\section{Complications and improvements}
\label{section:remarks} \todo{Bas, er staat bij al het werk van anderen een referentie}
In this section, we quickly make some further remarks on how or when to apply the theory we have built up in previous section to real cases.
\begin{itemize} 
\item Even if $r \geq g$, it might still be possible that $\overline{J(\Z)}$ coincidentally has dimension $r'$ smaller than $r$ (for example, if $r > g$); then one can find a rank $r'$ subgroup of $J(\Z)$ with the same closure inside $J(\Z_p)$, and do computations in this subgroup; we can expect this to work if $r' < g$. We can even use this to improve our calculations if we already have $r < g$, but $r'$ is even smaller. On the other hand, if $J(\Z)$ is dense in $J(\Z_p)$, there is no way to apply Chabauty's method to the curve.
\item Note that in all of the discussion of the previous section, we are calculating upper bounds for the cardinality of $C(\Z_p)_P \cap \overline{J(\Z)_t}$. It can and does however happen that this set is strictly bigger than $C(\Z)_P$, as shown in \cite{Balakrishnan19}, where Coleman-Chabauty is used on a database of 16997 curves of genus $3$ with Mordell-Weil rank of the Jacobian equal to $1$. They treat several possible cases and examples where the set is bigger. One of the main families they have found can be given in the following way: let $Q$ be a $K$-point $Q$ where $K$ is a quadratic number field where $p$ splits. Embedding $K$ into $\Q_p$, assume $Q-b$ lies in $J(\Z_p)_0$ and is torsion of order prime to $p$ in $J(\Z_p)_0/J(\Z)_0$. Then, as $J(\Z_p)$ is isomorphic as continous group to $\Z_p^g$, the point $Q-b$ lies in $\overline{J(\Z)_0}$, and clearly also in $C(\Z_p)$.\todo{Bas, even over hebben}
\item An important part of doing these calculations, is working with the Mordell-Weil group $J(\Z)$. Finding generators of $J(\Z)$ is as of yet a computationally difficult problem, and one may need to assume the BSD conjecture to even find the rank $r$. However, note that one does not need to find all of $J(\Z)$; if we can generate a subgroup of $J(\Z)$ of index finite and coprime to $p|J(\F_p)|$, it will have the same closure in $J(\Z_p)$. For this, we only need to generate sufficiently many independent points of the Jacobian, and saturate subgroups with respect to some primes. Both of these tasks are easier than finding all of $J(\Z)$.
\item Even this method does not work for a certain prime $p$, and neither for some other primes we tried, we can still use parts of the information we have learned. For example, maybe the method tells us that we have found all points of $J(\Z)$ stemming from $C(\Z)$ except those in a certain fibre of the map $J(\Z) \to J(\F_p)$. We can then aggregate this information for different primes, by looking at the maps $J(\Z) \to \prod_{p \in S} J(\F_p)$ for some set $S$ of primes of good reduction. See for example \cite{stoll10} for an introduction to the Mordell-Weil sieve.
\end{itemize}

@@

Code for the calculations in this section can be found in \cite{spelier2020}.
